---
title: "Use the Standardized Bradford Distribution in R with `bradfordr`"
author: "Victor Dogo"
date: "Last updated in `r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use the Standardized Bradford Distribution in R with `bradfordr`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bradfordr)
```

The aim of `{bradfordr}`, initially, was to simply generate random numbers from the Standardized Bradford Distribution. It uses the **Inverse Sampling Method** to generate numbers from it:

## The Inverse Transform Sampling Method

The Inverse Transform Sampling Method is widely used to generate pseudo-random variables in cases where a specific distribution isn't available. It can be applied both computationally and analytically. 

In short, this method consists in the following steps:

1. A number $u$ between 0 and 1 is sampled from the $Uniform(0,1)$ distribution and interpreted as a **probability**;
2. We take the distribution's **CDF** and equalize it to $u$. In our case, we have:

$$
u = F(x;c) = \frac{log(1 + cx)}{log(1+c)}
$$

3. The expression obtained is inverted; in other words, we manipulate the equation so that $x$ is isolated. In our case, we have:

$$
x = F^{-1}(u;c) = \frac{(1+c)^u - 1}{c}
$$

4. That's it! The $x$ value obtained is a pseudo-random number generated from the distribution of interest.

A more in depth and technical definition can be seen in [Wikipedia](https://en.wikipedia.org/wiki/Inverse_transform_sampling).

## Generating Numbers

It's as easy as 1-2-3: just call the `rbradford` function with sample size `n` and parameter `c` (defaults to `c=5`):

```{r}

data <- rbradford(1000, c=10)

head(data)

```

We can easily plot this data to understand how it behaves:

```{r}

hist(data)

```

With greater `c` values, comes a longer tail:

```{r}

data <- rbradford(1000, c=1000)

hist(data)

```

These differences can be seen better with the help of the other functions. It's also possible to verify that graphically:

## Verify distribution behavior graphically

It's useful, for us, to visualize how the distribution's curve changes regarding the possible values of the `c` parameter. This can be done with `pbradford` (which defines the Cumulative Distribution Function), `dbradford` (which defines the Probability Density Function) and `qbradford` (which defines the Quantile Function). 

In this vignette, I'll give an example with `pbradford`, which is the most useful function of the three mentioned above.

```{r}
# Behavior of cdf for c = 1, 10 and 100

x <- seq(0,1,0.01)

# Defining different values for the cdf based on c parameter
y_c1 <- pbradford(x,c=1)
y_c10 <- pbradford(x,c=10)
y_c100 <- pbradford(x,c=100)

# Plotting

plot(x,y_c1,type="l",col="red",main="CDF of Bradford Dist.",ylab="Fx")
lines(x,y_c10,col="blue")
lines(x,y_c100,col="green")
legend("bottomright",legend=c("c=1","c=10","c=100"),
       lty = c(1,1,1),
       col=c("red","blue","green"))


```

Here, we can see that, as `c` gets bigger, so does the weight to the right of the cdf. In other words, this means that small values of `x` applied to $F(x;c) = P(X \leq x)$ tend to give us bigger probabilities as `c` grows.

This is true to all other functions as well: the size of the parameter `c` influences how much weight smaller values of $x$ have. 

## Remarks

This vignette is a WIP. Any suggestions are welcome here. I'm always finding better ways to write everything in this package, so comments help me a lot.
