---
title: "Generating Random Numbers from the Standardized Bradford Distribution with `bradfordr`"
author: "Victor Dogo"
date: "Last updated in `r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating Random Numbers from the Standardized Bradford Distribution with `bradfordr`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bradfordr)
```

The aim of `{bradfordr}`, initially, was to simply generate random numbers from the Standardized Bradford Distribution. It uses the **Inverse Sampling Method** to generate numbers from it:

## The Inverse Transform Sampling Method

The Inverse Transform Sampling Method is widely used to generate pseudo-random variables in cases where a specific distribution isn't available. It can be applied both computationally and analytically. 

In short, this method consists in the following steps:

1. A number $u$ between 0 and 1 is sampled from the $Uniform(0,1)$ distribution and interpreted as a **probability**;
2. We take the distribution's **CDF** and equalize it to $u$. In our case, we have:

$$
u = F(x;c) = \frac{log(1 + cx)}{log(1+c)}
$$

3. The expression obtained is inverted; in other words, we manipulate the equation so that $x$ is isolated. In our case, we have:

$$
x = F^{-1}(u;c) = \frac{(1+c)^u - 1}{c}
$$

4. That's it! The $x$ value obtained is a pseudo-random number generated from the distribution of interest.

A more in depth and technical definition can be seen in [Wikipedia](https://en.wikipedia.org/wiki/Inverse_transform_sampling).

## Generating Numbers

It's as easy as 1-2-3: just call the `rbradford` function with sample size `n` and parameter `c` (defaults to `c=5`):

```{r}

data <- rbradford(1000, c=10)

head(data)

```

We can easily plot this data to understand how it behaves:

```{r}

hist(data)

```

With greater `c` values, comes a longer tail:

```{r}

data <- rbradford(1000, c=1000)

hist(data)

```

These differences can be seen better with the help of the other functions. It's also possible to verify that graphically:

##

